---
title: "3. Create KEGG profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Create KEGG profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r include=FALSE}
library(kableExtra)
library(tidyverse)
library(readxl)
library(DT)
```

Rbims also includes functions to read and explore [KEGG](https://www.genome.jp/kegg/) annotations. To learn how to to this; 

Make sure the library is loaded 

```{r setup, eval=F}
library(rbims)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all()
```

## `read_ko()`: Parse KEGG Outputs

The [`read_ko`](https://mirnavazquez.github.io/RbiMs/reference/read_ko.html). function is designed to parse and format raw output files from the InterProScan annotation tool.


```{r eval=FALSE}
?read_ko
```
 

### Input Requirements

- File Path: Provide the path to the directory containing your KEGG files (outputs from the KofamKOALA/KofamScan annotation). Even if its just one, the path needs to be explicitly defined for the directory only. 

- File Extension: The function specifically looks for files ending in `*.txt`.

### Data Structure

The processed input contains 6 key columns:

* gene name: the unique identifier for the sequence that was annotated.                    
* KO: KEGG ortholog identifier. The "ID number" for the functional group.

* thrshld: the minimum score requirement set by the annotation tool to consider a match "real."

* score: the bit-score from the sequence alignment.

* E-value: the number of hits one can "expect" to see by chance when searching a database of a particular size.

* KO definition: description of the functional group. 


If you want to follow this example, you can download the files from [here](https://github.com/KarlaaPLM/rbims_hydrocarbon/tree/main/Functional%20annotation/KEGG).

```{r eval=FALSE, warning=FALSE, include=FALSE}
ko_bin_table <-read_ko(data_kofam = "/Users/karliz/Desktop/Estancia_ICMyL_Mirna/Hidrocarburos/KEGG_hidro",
                                   write = F)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
ko_bin_table <- read_tsv("~/Documents/Rbims/inst/extdata/ko_bin_table.tsv", col_names = T)
```


```{r eval=FALSE}
ko_bin_table<-read_ko(data_kofam ="../test/results/02.kofam")
```

The read_ko function will create a table that contains the abundance of each KO within each bin.

```{r eval = F}
head(ko_bin_table)
```


```{r echo=FALSE}
ko_bin_table %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. KEGG Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(ko_bin_table), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


## Map to the KEGG database

The function [mapping_ko](https://mirnavazquez.github.io/RbiMs/reference/mapping_ko.html) can now be used to map the KO and their abundance to the rest of the features of KEGG and rbims database.


```{r message=FALSE, warning=FALSE, include=FALSE}
ko_bin_mapp <- read_tsv("~/Documents/Rbims/inst/extdata/ko_bin_mapp.tsv", col_names = T)
```

```{r eval=FALSE}
ko_bin_mapp<-mapping_ko(ko_bin_table)
```

```{r eval=FALSE}
head(ko_bin_mapp)
```


```{r echo=FALSE}
ko_bin_mapp %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. KEGG Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(ko_bin_mapp), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


You can export this profile like this:

```{r, eval=FALSE}
write.table(ko_bin_mapp, "ko_bin_mapp.tsv", 
            quote = F, 
            sep = "\t", 
            row.names = F, 
            col.names = T)
```

Or setting write  `write = T` in the function `read_ko()`



