---
title: "5. Create dbCAN profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Create dbCAN profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
library(kableExtra)
library(tidyverse)
library(readxl)
library(DT)
```

Rbims also includes functions to read and explore [dbCAN](https://bcb.unl.edu/dbCAN2/) annotations. To learn how to to this; 

First, load the rbims package.

```{r, eval=FALSE}
library(rbims)
```


```{r, echo=FALSE, warning=F, message=FALSE}
devtools::load_all()
```

## `read_dbcan3()`: Parse dbCAN3 Outputs
The `read_dbcan3` function is designed to parse and format raw output files from the dbCAN3 annotation tool.

### Input Requirements

- File Path: Provide the path to the directory containing your dbCAN files.

- File Extension: The function specifically looks for files ending in `*.overview.txt`.

### Data Structure

The processed input contains 6 key columns:

- Bin Name: Identifier of the bin/genome.

- HMMER / Hotpep / DIAMOND: Genes identified by each specific algorithm.

- SignalP: Indicates if a signal peptide was detected.

- #ofTools: Total count of algorithms that identified the gene.

If you want to follow this example, you can download the raw data [here](https://github.com/mirnavazquez/RbiMs/tree/main/inst/extdata/dbCAN).

To obtain a wide output, the argument `profile = T`.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
dbcan_profile_T <-read_dbcan3(dbcan_path = "/Users/karliz/Desktop/Estancia_ICMyL_Mirna/Hidrocarburos/dbCAN_hidro",  profile = T, write = F)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
dbcan_profile_T <- read_tsv("~/Documents/Rbims/inst/extdata/dbcan_profile_T.tsv", col_names = T)
```

```{r eval=FALSE, warning=FALSE}
dbcan_profile_T <-read_dbcan3(dbcan_path = "../test/results/03.dbcan",  
                              profile = T, 
                              write = F)
```

Some extra information is recovered from the input files, such as:

- Total number of genes.

- Remaining genes after the the filtered.

- Number of genes that have signals and passed the filtered. 

```{r eval = FALSE}
head(dbcan_profile_T)
```


```{r echo=FALSE}
dbcan_profile_T %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. dbCAN Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(dbcan_profile_T), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


Or print a long table `profile = F`.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
dbcan_profile_F <-read_dbcan3(dbcan_path = "/Users/karliz/Desktop/Estancia_ICMyL_Mirna/Hidrocarburos/dbCAN",  profile = F, write = F)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
dbcan_profile_F <- read_tsv("~/Documents/Rbims/inst/extdata/dbcan_profile_F.tsv", col_names = T)
```


```{r eval=FALSE, warning=FALSE}
dbcan_profile_F <-read_dbcan3(dbcan_path = "../test/results/03.dbcan",  
                              profile = F, 
                              write = F)
```


```{r eval = FALSE}
head(dbcan_profile_F)
```


```{r echo=FALSE}
dbcan_profile_F %>%
  head(15) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 2. dbCAN Profile Overview (`profile = F`)</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = TRUE, 
                position = "left") %>%
  column_spec(1:ncol(dbcan_profile_F), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


You can export this profile with: 

```{r, eval=FALSE}
write.table(dbcan_profile_T, "dbcan.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Or setting write  `write = T` in the function `read_dbcan3()`
