---
title: "9. Explore InterProScan profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{9. Explore InterProScan profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=7
)
```

```{r include=FALSE}
library(kableExtra)
library(tidyverse)
library(readxl)
library(DT)
```


First, make sure the library is loaded 

```{r setup, eval=F}
library(rbims)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all()
```
If you want to follow this example, you can download the files from [here](https://github.com/KarlaaPLM/rbims_hydrocarbon/tree/main/Functional%20annotation/InterProScan) file. 

## Now let's load the long Interpro profile using the "Pfam" database:

```{r message=FALSE, warning=FALSE, include=FALSE}
pfam_profile <- read_tsv("~/Documents/Rbims/inst/extdata/interpro_profile_T.tsv", col_names = T)
```

```{r eval=FALSE, warning=FALSE}
pfam_profile <-read_interpro(data_interpro = "../inst/extdata/Interpro",
                             database = "Pfam",
                             profile = T, 
                             write = F)
```

```{r eval = FALSE}
head(pfam_profile)
```

```{r echo=FALSE}
pfam_profile %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. InterproScan Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(pfam_profile), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


## To explore the *Interpro* profiles we can use the following function:

- `get_subset_pca`: which performs a PCA analysis to reduce the dimensions of the contributing families, resulting in a new table with the most represented Pfam families (for this example) of the bins/genome samples. 

### Some considerations:
  
  * To better visualize the results we recommend the loading of metadata, which essentially could include:

  * The clean names of the bins

  * Type of sampling

  * Environment 

  * Hierarchical order
  
```{r}
metadata <- read_excel("../inst/extdata/metadata_SIPH.xlsx")
```

```{r eval=FALSE}
head(metadata)
```

```{r echo=FALSE}
metadata %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 2. Metadata overview </span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(metadata), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```

```{r echo=TRUE}
#Extra commands to clean the bin names 

lookup <- setNames(metadata$Bin_name, metadata$Clean_name)

pfam_profile_renamed <- pfam_profile %>% 
  rename(!!!lookup) 
```


```{r message=FALSE, warning=FALSE}
important_PFAMs<-get_subset_pca(tibble_rbims=pfam_profile_renamed,
                                cos2_val=0.96,
                                analysis="Pfam")
```

```{r eval=FALSE}
head(important_PFAMs)
```

```{r include=FALSE}
important_PFAMs %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 3. Most contributing InterproScan families </span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(important_PFAMs), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```

## To visualize the *InterproScan* subset we can use the following functions:

- [plot_heatmap](https://mirnavazquez.github.io/RbiMs/reference/plot_heatmap.html) 
- [plot_bubble](https://mirnavazquez.github.io/RbiMs/reference/plot_bubble.html) 

Use `plot_heatmap` with the argument `distance = T` to visualize the clustering of the bins/genome samples based on the InterproScan domains. 
 
```{r fig.cap="Figure 1. Clustering of InterproScan families across bins.", warning=FALSE, fig.width=9.7, fig.height=9, fig.align='center'}

plot_heatmap(important_PFAMs, 
             y_axis=Pfam, 
             analysis = "INTERPRO", 
             distance = T)

```

Use `plot_heatmap` with the argument `distance = F` to visualize the abundance of InterproScan domains across the bins/genome samples. 

```{r fig.cap="Figure 2. Presence/Absence of InterproScan families across bins.", warning=FALSE, fig.width=9, fig.height=6, fig.align='center'}

plot_heatmap(important_PFAMs, 
             y_axis=Pfam, 
             analysis = "INTERPRO", 
             distance = F)
```

```{r include=FALSE}
metadata_renamed <- read_excel("../inst/extdata/metadata_SIPH_renamed.xlsx")

#Calc argument to show presence/absence
ordered <- c("g_Flavobacterium_5m_16",
             "g_Flavobacterium_5m_26",
             "g_Henriciella_5m_15",
             "g_Hyphomonas_5m_32",
             "g_Hyphomonas_5m_33",
             "g_Celeribacter_5m_10",
             "g_Celeribacter_5m_0",
             "s_Planktomarina_temperata_5m_1",
             "s_Lentibacter_algarum_5m_13",
             "s_Lentibacter_algarum_5m_7",
             "g_Tateyamaria_5m_25", 
             "g_Tateyamaria_5m_8",
             "o_Pseudomonadales_5m_2",
             "s_Thalassolituus_oleivorans_5m_5",
             "s_Thalassolituus_oleivorans_5m_19",
             "g_Pseudophaeobacter_5m_3",
             "g_Pseudophaeobacter_5m_37",
             "g_Glaciecola_5m_9",
             "s_Alcanivorax_jadensis_5m_11",
             "g_Alcanivorax_5m_25",
             "g_Alcanivorax_5m_18",
             "s_Marinobacter_salarius_5m_14",
             "s_Marinobacter_salarius_5m_10",
             "g_Oleibacter_5m_18",
             "g_Pseudophaeobacter_700m_8",
             "g_Pseudophaeobacter_700m_13",
             "g_Glaciecola_700m_16",
             "g_Glaciecola_700m_18",
             "g_Alcanivorax_700m_20",
             "s_Marinobacter_salarius_700m_3",
             "s_Marinobacter_salarius_700m_24",
             "g_Oleibacter_700m_21",
             "g_Oleibacter_700m_15",
             "g_Olleya_700m_17",
             "g_Olleya_700m_14",
             "g_Dokdonia_700m_23",
             "g_Dokdonia_700m_12",
             "g_Paracoccus_700m_9",
             "g_Paracoccus_700m_1",
             "g_Sulfitobacter_700m_0",
             "g_Alteromonas_700m_22",
             "g_Alteromonas_700m_2")
```

For `plot_bubble`, the user has two options for the argument `calc`:

- Binary

```{r fig.align='center', fig.cap="Figure 3. Presence of InterproScan families across bins.", fig.width=9.8, fig.height=6, warning=FALSE}

plot_bubble(important_PFAMs, 
            y_axis=Pfam, 
            x_axis=Bin_name, 
            calc = "Binary",
            analysis = "INTERPRO", 
            data_experiment = metadata_renamed, 
            color_character = Depth,
            order_bins = ordered,
            y_labs = "Most contributing InterproScan families",
            text_y = 10,
            x_labs = "Bins",
            text_x = 9)
```
 
 Or 
 
 - Abundance 
 
```{r fig.cap="Figure 4. Abundance of InterproScan families across bins.", warning=FALSE, fig.width=9.8, fig.height=6, fig.align='center'}


plot_bubble(important_PFAMs, 
            y_axis=Pfam, 
            x_axis=Bin_name, 
            calc = "Abundance",
            analysis = "INTERPRO", 
            data_experiment = metadata_renamed, 
            color_character = Class,
            y_labs = "Most contributing InterproScan families",
            text_y = 10,
            x_labs = "Bins",
            text_x = 9,
            range_size = c(1,15))

```

# Exploring other databases within InterproScan

Because the annotation with InterproScan recovers data from different databases, rbims can also explore not only Pfams, but also TigerFams, Superamilies or InterproScan ID's. 
First, we will reload the InterProScan output in a wide format and extract the PFAM abundance information.

## Let's load the long Interpro profile using the "INTERPRO" database:

```{r message=FALSE, warning=FALSE, include=FALSE}
interpro_profile <- read_tsv("~/Documents/Rbims/inst/extdata/Interpro_T.tsv", col_names = T)
```


```{r eval=FALSE, warning=FALSE}
interpro_profile <-read_interpro(data_interpro ="../inst/extdata/Interpro",
                                 profile = T, 
                                 database = "INTERPRO",
                                 write = F)
```

```{r eval = FALSE}
head(interpro_profile)
```

```{r echo=FALSE}
interpro_profile %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. dbCAN Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(interpro_profile), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


## We are going to look for the InterProScan IDs that conform the `DNA topoisomerase 1`. To do this, we will create a vector of the IDs associated to that enzyme.

```{r}
DNA_topoisomerase_1<-c("IPR013497", "IPR023406", "IPR013824")
```

## With the function [get_subset_pathway](https://mirnavazquez.github.io/RbiMs/reference/get_subset_pathway.html) we can create a subset of the INTERPRO table.

```{r}
DNA_topo_INTERPRO<-get_subset_pathway(interpro_profile,
                                      type_of_interest_feature=INTERPRO,
                                      interest_feature=DNA_topoisomerase_1)
```

```{r eval =FALSE}
head(DNA_topo_INTERPRO)
```

```{r echo=FALSE}
DNA_topo_INTERPRO %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 4. InterproScan Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(DNA_topo_INTERPRO), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```

## To visualize it, we can create a bubble plot to explore the distribution of these enzymes across the bins.

```{r fig.cap="Figure 5. Abundance of InterproScan families conforming the DNA topoisomerase pathway across bins.", warning=FALSE, fig.width=9.8, fig.height=6, fig.align='center'}

plot_bubble(DNA_topo_INTERPRO, 
            y_axis=INTERPRO,
            x_axis=Bin_name,
            calc = "Abundance",
            analysis = "INTERPRO", 
            data_experiment = metadata, 
            color_character = Depth)
```

<!-- # Example with KEGG database -->

<!-- First, I will read the InterProScan output in a long format and extract the KEGG information. When you use the `KEGG` option, the profile option is disabled.  -->

<!-- ```{r, eval=FALSE} -->
<!-- interpro_KEGG_long<-read_interpro(data_interpro = "Interpro_test.tsv", database="KEGG") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- head(interpro_KEGG_long) -->
<!-- ``` -->

<!-- ## Mapping INTERPRO to KEGG database -->

<!-- We can use the [mapping_ko](https://mirnavazquez.github.io/RbiMs/reference/mapping_ko.html) function here, to get the extended KEGG table. -->

<!-- ```{r, eval=FALSE} -->
<!-- interpro_map<-mapping_ko(tibble_interpro = interpro_KEGG_long) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- head(interpro_map) -->
<!-- ``` -->

<!-- We can plot all the KOs and the Modules to which they belong. An important thing here is that we will set `analysis = "KEGG"` despite this workflow started with the InterProScan output in analysis.  -->

<!-- ```{r, fig.alt="Heatmap with metadata"} -->
<!-- plot_heatmap(tibble_ko=interpro_map, -->
<!--              data_experiment = metadata, -->
<!--              y_axis=KO, -->
<!--              order_y = Module, -->
<!--              order_x = Sample_site, -->
<!--              split_y = TRUE, -->
<!--              analysis = "KEGG", -->
<!--              calc="Percentage") -->
<!-- ``` -->












