---
title: "11. Explore MEROPS profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{11. Explore_MEROPS_profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all()
```


```{r include=FALSE}
library(kableExtra)
library(tidyverse)
library(readxl)
library(DT)
```

# How to explore the MEROPS profile? 

First, make sure the library is loaded 

```{r setup, eval=F}
library(rbims)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all()
```

## Now let's load the MEROPS profile:


```{r message=FALSE, warning=FALSE, include=FALSE}
merops_profile_T <- read_tsv("~/Documents/Rbims/inst/extdata/merops_profile_T.tsv", col_names = T)
```

```{r eval=FALSE, warning=FALSE}
merops_profile_T <- read_merops(merops_path = "../inst/extdata/MEROPS", profile = T, write = F )
```

```{r eval = FALSE}
head(merops_profile_T)
```

```{r echo=FALSE}
merops_profile_T %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 1. MEROPS Profile Overview ( `profile = T` )</span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(merops_profile_T), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```

## To explore the *MEROPS* profiles we can use the following function:

- `get_subset_pca`: which performs a PCA analysis to reduce the dimensions of the contributing families, resulting in a new table with the most represented MEROPS families of the bins/genome samples. 

### Some considerations:

* To better visualize the results we recommend the loading of metadata, which essentially could include:

  * The clean names of the bins

  * Type of sampling

  * Environment 

  * Hierarchical order

```{r}
metadata <- read_excel("../inst/extdata/metadata_SIPH.xlsx")
```

```{r eval=FALSE}
head(metadata)
```

```{r echo=FALSE}
metadata %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 2. Metadata overview </span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(metadata), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```


```{r echo=TRUE}
#Extra commands to clean the bin names 

lookup <- setNames(metadata$Bin_name, metadata$Clean_name)

MEROPS_Hidro_T_renamed <- merops_profile_T %>% 
  rename(!!!lookup) 

```

### Running `get_subset_pca`: 

```{r warning=FALSE}
important_merops_FAMs<-get_subset_pca(tibble_rbims=MEROPS_Hidro_T_renamed, 
                                     cos2_val=0.85,
                                     analysis="MEROPS")
```

```{r eval=FALSE}
head(important_merops_FAMs)
```

```{r include=FALSE}
important_merops_FAMs %>%
  head(10) %>%
  knitr::kable(
    caption = "<span style='display: block; text-align: left; width: 100%; font-weight: bold;'>Table 3. Most contributing MEROPS families </span>",
    align = "c",
    format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"),
                full_width = FALSE, 
                position = "left") %>%
  column_spec(1:ncol(important_merops_FAMs), border_right = "1px solid #ddd", border_left = "1px solid #ddd", extra_css = "text-align: center;") %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2") %>%
  scroll_box(height = "300px", width = "100%", fixed_thead = TRUE)
```

## To visualize the *MEROPS* subset we can use the following functions:

- [plot_heatmap](https://mirnavazquez.github.io/RbiMs/reference/plot_heatmap.html) 
- [plot_bubble](https://mirnavazquez.github.io/RbiMs/reference/plot_bubble.html) 

Use `plot_heatmap` with the argument `distance = T` to visualize the clustering of the bins/genome samples based on the MEROPS domains. 
 
```{r fig.cap="Figure 1. Clustering of MEROPS families across bins.", warning=FALSE, fig.width=9.7, fig.height=11, fig.align='center'}

plot_heatmap(important_merops_FAMs, 
             y_axis=MEROPS_family, 
             analysis = "MEROPS", 
             distance = T)

```

Use `plot_heatmap` with the argument `distance = F` to visualize the presence/absence of MEROPS domains across the bins/genome samples. 

```{r fig.cap="Figure 2. Presence/Absence of MEROPS families across bins.", warning=FALSE, fig.width=9, fig.height=11, fig.align='center'}

plot_heatmap(important_merops_FAMs, 
             y_axis=MEROPS_family, 
             analysis = "MEROPS", 
             distance = F)
```


```{r include=FALSE}
metadata_renamed <- read_excel("../inst/extdata/metadata_SIPH_renamed.xlsx")

#Calc argument to show presence/absence
ordered <- c("g_Flavobacterium_5m_16",
             "g_Flavobacterium_5m_26",
             "g_Henriciella_5m_15",
             "g_Hyphomonas_5m_32",
             "g_Hyphomonas_5m_33",
             "g_Celeribacter_5m_10",
             "g_Celeribacter_5m_0",
             "s_Planktomarina_temperata_5m_1",
             "s_Lentibacter_algarum_5m_13",
             "s_Lentibacter_algarum_5m_7",
             "g_Tateyamaria_5m_25", 
             "g_Tateyamaria_5m_8",
             "o_Pseudomonadales_5m_2",
             "s_Thalassolituus_oleivorans_5m_5",
             "s_Thalassolituus_oleivorans_5m_19",
             "g_Pseudophaeobacter_5m_3",
             "g_Pseudophaeobacter_5m_37",
             "g_Glaciecola_5m_9",
             "s_Alcanivorax_jadensis_5m_11",
             "g_Alcanivorax_5m_25",
             "g_Alcanivorax_5m_18",
             "s_Marinobacter_salarius_5m_14",
             "s_Marinobacter_salarius_5m_10",
             "g_Oleibacter_5m_18",
             "g_Pseudophaeobacter_700m_8",
             "g_Pseudophaeobacter_700m_13",
             "g_Glaciecola_700m_16",
             "g_Glaciecola_700m_18",
             "g_Alcanivorax_700m_20",
             "s_Marinobacter_salarius_700m_3",
             "s_Marinobacter_salarius_700m_24",
             "g_Oleibacter_700m_21",
             "g_Oleibacter_700m_15",
             "g_Olleya_700m_17",
             "g_Olleya_700m_14",
             "g_Dokdonia_700m_23",
             "g_Dokdonia_700m_12",
             "g_Paracoccus_700m_9",
             "g_Paracoccus_700m_1",
             "g_Sulfitobacter_700m_0",
             "g_Alteromonas_700m_22",
             "g_Alteromonas_700m_2")
```

For `plot_bubble`, the user has two options for the argument `calc`:

- Binary

```{r fig.align='center', fig.cap="Figure 3. Presence of MEROPS families across bins.", fig.width=9.8, fig.height=11, warning=FALSE}

plot_bubble(important_merops_FAMs, 
            y_axis=MEROPS_family, 
            x_axis=Bin_name, 
            calc = "Binary",
            analysis = "MEROPS", 
            data_experiment = metadata_renamed, 
            color_character = Depth,
            order_bins = ordered,
            y_labs = "Most contributing MEROPS families",
            text_y = 10,
            x_labs = "Bins",
            text_x = 9)
```
 
 Or 
 
 - Abundance 
 
```{r fig.cap="Figure 4. Abundance of MEROPS families across bins.", warning=FALSE, fig.width=9.8, fig.height=7, fig.align='center'}


plot_bubble(important_merops_FAMs, 
            y_axis=domain_name, 
            x_axis=Bin_name, 
            calc = "Abundance",
            analysis = "MEROPS", 
            data_experiment = metadata_renamed, 
            color_character = Class,
            y_labs = "Most contributing MEROPS families",
            text_y = 10,
            x_labs = "Bins",
            text_x = 9,
            range_size = c(1,5))

```

