---
title: "Create Interpro profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create Interpro profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(rbims)
```

A common practice to get other annotations of genomes and reconstructed genomes is to use [InterProScan](https://interproscan-docs.readthedocs.io/en/latest/). Here are some functions to explore that information.

The function to use that information is `read_interpro`. This functio can parse the information of the PFAM, INTEPRO and KEGG ids. The KEGG analysis is just possible if [InterProScan](https://interproscan-docs.readthedocs.io/en/latest/) was run with the -pa option. As well tow output options are possible a wide profile or a long table. 


- The `database` argumen indicate which database will be parse, in this example I will explore the PFAM output. 

- The output format is control with the `profile` argument. When **profile = T** then a wide output is obtained.

```{r, eval=FALSE}
interpro_pfam_profile<-read_interpro("Interpro_test.tsv", database="PFAM", profile = T)
```

```{r}
head(interpro_pfam_profile)
```

Or print a long table **profile = F**. 

```{r, eval=FALSE}
interpro_pfam_long<-read_interpro("Interpro_test.tsv", database="PFAM", profile = F)
```

```{r}
head(interpro_pfam_long)
```








