---
title: "Create KEGG profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create KEGG profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, load the rbims package.

```{r setup}
library(rbims)
```

The second thing to do would be to read the [KofamKOALA/KofamScan](https://github.com/mirnavazquez/RbiMs/blob/main/inst/extdata/KEGG_bins.txt) output. 

## Read the KEGG results

The function that rbims uses to read the raw output from the KofamKOALA/KofamScan data and make a table is [read_ko](https://mirnavazquez.github.io/RbiMs/reference/read_ko.html).

```{r eval=FALSE}
?read_ko
```

`rbims` contains a test dataset that allow us to test this function. This dataset is saved in objects: `ko_bin_table`, `ko_bin_mapp`, and `metadata`. First download the [KofamKOALA/KofamScan](https://github.com/mirnavazquez/RbiMs/blob/main/inst/extdata/KEGG_bins.txt) example file. It is recommended that you save this file in its own folder since this function will read all the text files in your path and concatenate them. An example of a path input is shown below: 

```{r eval=FALSE}
ko_bin_table<-read_ko(data_kofam ="C:/Users/Bins")
```

The read_ko function will create a table that contains the abundance of each KO within each bin.  

```{r}
head(ko_bin_table)
```

## Map to the KEGG database

The function [mapping_ko](https://mirnavazquez.github.io/RbiMs/reference/mapping_ko.html) can now be used to map the KO and their abundance to the rest of the features of KEGG and rbims database.

```{r eval=FALSE}
ko_bin_mapp<-mapping_ko(ko_bin_table)
```

```{r}
head(ko_bin_mapp)
```

You can export this to a table like this:

```{r}
write.table(ko_bin_mapp, "KEGG_mapped.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```



