---
title: "Create KEGG profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create KEGG profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(rbims)
```

The first thing to do would be read the KofamKOALA/KofamScan output and the
metadata file. The metadata file is a tab-separated file containing the
name of your bins and any extra information you would like to use for
visualization.

The function that rbims use to read the input table is `read_ko`. And it reads the raw output from the  KofamKOALA/KofamScan programs.

```{r eval=FALSE}
?read_ko
```

`rbims` contains external data that allow us to test this function. You can download the file from here (https://github.com/mirnavazquez/RbiMs/blob/main/inst/extdata/KEGG_bins.txt) and try. 

```{r eval=FALSE}
ko_bin_table<-read_ko("KEGG_bins.txt")
```

The read_ko function will create a table that contains the abundance of each KO within each bin. 

```{r}
head(ko_bin_table)
```

The metadata file is a file where we have external information of each bin, like sample site or taxonomy. This information can be used in downstream analysis. An example of a metadata file is included in rbims.

To read the metadata we will use the function `read_excel` from the package `readxl`. Then you can download the file from here: https://github.com/mirnavazquez/RbiMs/blob/main/inst/extdata/metadata.xlsx, and try.

```{r, eval=FALSE}
metadata<-read_excel("metadata.xlsx")
```

## Map to the KEGG database

The function `mapping_ko` can be now used to map the KO and their abundance to the rest of the features of KEGG and rbims database.

```{r eval=FALSE}
ko_bin_mapp<-mapping_ko(ko_bin_table)
```


```{r}
head(ko_bin_mapp)
```


