---
title: "Create Interpro profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create Interpro profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A common practice to annotate the genes of a new genome or reconstructed genome is to use [InterProScan](https://interproscan-docs.readthedocs.io/en/latest/). Here are some functions to explore that information.

First, load the rbims package.

```{r setup}
library(rbims)
```

The function to use that information is `read_interpro`. This function can parse the information of the PFAM, INTERPRO, and KEGG ids. The KEGG analysis is just possible if [InterProScan](https://interproscan-docs.readthedocs.io/en/latest/) was run with the -pa option. Two output options are also possible: a wide profile, or a long table. 

- The `database` argument will parse the database. In this example, I will explore the PFAM output. 

- The output format is chosen with the `profile` argument. When **profile = T**, a wide output is obtained.

If you want to follow the example you can download the use rbims [test](https://github.com/mirnavazquez/RbiMs/blob/main/inst/extdata/Interpro_test.tsv) file. 

```{r, eval=FALSE}
interpro_pfam_profile<-read_interpro(data_interpro = "Interpro_test.tsv", database="Pfam", profile = T)
```

```{r}
head(interpro_pfam_profile)
```

Or print a long table **profile = F**. 

```{r, eval=FALSE}
interpro_pfam_long<-read_interpro("Interpro_test.tsv", database="Pfam", profile = F)
```

```{r}
head(interpro_pfam_long)
```

You can export this to a table like this:

```{r, eval=FALSE}
write.table(interpro_pfam_long, "Interpro.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```









