---
title: "Create picrust2 profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create picrust2 profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Create picrust2 profile

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Another tool to use in metabarcoding is  [picrust2](https://huttenhower.sph.harvard.edu/picrust/). Here is the function to explore these type of files:

First, load the rbims package.

```{r setup}
library(rbims)
```


The function to use that information is `read_picrust2`. This function can parse the information of the dbCAN3 files.

-   The `database` argument will parse the database. In this example, I will explore the KO output.

-   The output format is chosen with the `profile` argument. When **profile = T**, a wide output is obtained.

-   The `write` argument saves the formatted table generated in .tsv extension. When **write = F** gives you the output but not saves the table in your current directory. 

If you want to follow the example you can download the use rbims [test](https://github.com/Steph0522/Rbims_workflow/tree/main/inst/extdata_new) file.

```{r, eval=TRUE}
picrust2_profile<-read_picrust2(
data_picrust2 =  "../inst/extdata/pred_metagenome_unstrat.tsv",database = "KO",  profile = T, write = F)
```

```{r}
head(picrust2_profile)
```

Or print a long table **profile = F**.

```{r, eval=TRUE}
picrust2_profile<-read_picrust2(
data_picrust2 = "../inst/extdata/pred_metagenome_unstrat.tsv",  database = "KO",
                           profile = F, write = F)
```

```{r}
head(picrust2_profile)
```



You can export this to a table like this:

```{r, eval=FALSE}
write.table(picrust2_profile, "picrust2.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Or setting write  **write = T**. 