---
title: "Explore KEGG profile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explore KEGG profile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=7
)
```

```{r setup}
library(rbims)
```

The first thing to do would be read the kofamScan output and the
metadata file. The metadata file is a tab-separated file containing the
name of your bins and any extra information you would like to use for
visualization.


```{r, eval=FALSE}
ko_bin_table<-read_ko(data_kofam = "all.kofam.txt")
metadata <-read.table("metadata.txt", sep = "\t")
```

Then map the KO to the rest of the features of the KEGG database and
rbims database.

```{r, eval=FALSE}
ko_bin_mapp<-mapping_ko(ko_bin_table)
```

Let’s say that you are interested in the genes associated with the
biofilm formation in Vibrio Cholerae.

-   One first thing could be to create a vector containing the name of
    the KEGG pathway associated with the biofilm formation in Vibrio
    Cholerae.

```{r}
Biofilm_Vibrio<-c("map05111")
```

-   Now, let’s extract the profile associated with that metabolic
    pathway.

```{r}
library(tidyr)
Biofilm_Vibrio_subset<-ko_bin_mapp%>%
  drop_na(Pathway) %>%
  get_subset_pathway(Pathway, Biofilm_Vibrio) 
```

-   Now, let’s create a plot of presence and absence of the different KO
    associated to that pathway. Besides presence and absence, it is
    possible to plot abundance or the percentage of genes within certain
    pathways (See ?plot\_bubble, calc argument).
    
```{r}
plot_bubble(tibble_ko = Biofilm_Vibrio_subset,
            x_axis = Bin_name, 
            y_axis = KO,
            analysis="KEGG",
            data_experiment = metadata,
            calc="Binary",
            color_character = Order,
            range_size = c(1,10)) 
```

### Order axis

Let’s say that you want to order by bin names.

-   Create a vector containing the order.

```{r}
order_taxa<-c("Bin_1", "Bin_2", "Bin_10", "Bin_113", "Bin_12", "Bin_56")
```

-   Now plot, using the order\_bins argument.

```{r}
plot_bubble(tibble_ko = Biofilm_Vibrio_subset,
            data_experiment = metadata,
            x_axis = Bin_name,
            y_axis = Genes,
            analysis="KEGG",
            calc="Binary",
            order_bins=order_taxa,
            color_character=Genus,
            range_size = c(5,6))
```

You can also visualize the data with the a heatmap.

```{r}
plot_heatmap(tibble_ko=Biofilm_Vibrio_subset, 
             y_axis=Genes,
             analysis = "KEGG",
             calc="Binary")
```


